Determinism, Complexity, and Predictability in Computer Performance

### 2014-04-11 intro.tex
red text edits

As a corollary of the undecidability of the halting
problem~\cite{halting-problem}, however, no single forecasting schema
is ideal for all noise-free deterministic
signals~\cite{weigend-book}---let alone all real-world time-series
data sets.  This leads naturally to an important and challenging
question: given a complicated real-valued time series, does there
exist a forecast model that can utilize the information (if any) that
is being transmitted from past to future by the underlying generating
process \alert{in order to predict it}?  A first step in answering this question is to reliably
quantify where on the complexity spectrum a given time series falls; a
second step is to determine how complexity and predictability are
related.  These are the goals of this paper.

>>>>

As a corollary of the undecidability of the halting
problem~\cite{halting-problem}, however, no single forecasting schema
is ideal for all noise-free deterministic
signals~\cite{weigend-book}---let alone all real-world time-series
data sets.  This leads naturally to an important and challenging
question: given a complicated real-valued time series, does there
exist a forecast model that can leverage the information (if any) that
is being transmitted from past to future by the underlying generating
process?  A first step in answering this question is to reliably
quantify where on the complexity spectrum a given time series falls; a
second step is to determine how complexity and predictability are
related.  These are the goals of this paper.

### 2014-04-11 methods.tex
edit through the model section
red text edits

\col is a simple three-line C program that repeatedly initializes the
upper triangle of a 2048 $\times$ 2048 matrix in column-major order.
As mentioned in Section~\ref{sec:intro}, this simple program exhibits
surprisingly complicated behavior~\cite{mytkowicz09}.  A time series
of the processor performance during the execution this program is
shown in Figure~\ref{fig:col-ts}.
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/colFullTS}
    \caption{The instructions executed per CPU clock cycle (IPC) as
      the \col program runs.  Each point is the average IPC in a
      100,000 instruction period. Figure~\ref{fig:col-ipc} is a
      closeup of the first part of this signal.  }
    \label{fig:col-ts}

>>>>

\col is a simple C program that repeatedly initializes the upper
triangle of a 2048 $\times$ 2048 matrix in column-major order.

{\color{red}Joshua, can you please insert code here?}

\noindent As mentioned in Section~\ref{sec:intro}, this simple program
exhibits surprisingly complicated behavior~\cite{mytkowicz09}.  A time
series of the processor performance during the execution this program
is shown in Figure~\ref{fig:col-ts}. \alert{would it be useful to
  include said 3-line loop?}
\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/colFullTS}
    \caption{The instructions executed per CPU clock cycle (IPC) as
      the \col program runs.  Each point is the average IPC in a
      100,000 instruction period. Figure~\ref{fig:col-ipc} is a
      closeup of the first part of this signal.  }
    \label{fig:col-ts}
